---
title: "Assignment5"
author: "Holly Finertie - HF2379"
date: 'Due: 2/25/2020'
output:
  word_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(caret)
library(glmnet)
library(modelr)
library(mgcv)

set.seed(100)
```

Goal: You want to predict current alcohol consumption but it is expensive and time-consuming to administer all of the behavioral testing that produces the personality scores. You will conduct a reproducible analysis to build and test classification models using regularized logistic regression and traditional logistic regression.

### Data Import: Cleaning and Training/Testing Data Set

I imported the data, cleaned the variable names, and converted our outcome of interest to a factor variable named 'alc_outcome'. Then I created training and testing data sets with a 70/30 split. 

```{r, import}
alc = read_csv("./data/alcohol_use.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    alc_outcome = as.factor(alc_consumption)
  )

head(alc)


train_alc = alc %>% sample_frac(.7)
test_alc = anti_join(alc, train_alc, by = 'x1')
```

